Gut-resident microorganisms and their genes are associated with cognition and neuroanatomy in children

### 2023-03-31 manuscript/main.tex
vanja comments 2

%%%%%%%%%%TO FORMAT A PREPRINT FOR A PCI%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% figures
\usepackage{wrapfig}
\usepackage{lscape}
\usepackage{rotating}
\usepackage{graphicx}
\graphicspath{assets}
\usepackage{float}

@@

\begin{table}[!h]
    \begin{center}
    \begin{tabular}{|r|r|r|r|r|r|r|}
      \hline\hline
      \textbf{Age group} & \multicolumn{2}{c}{\textbf{0 to 120 months}} & \multicolumn{2}{c}{\textbf{0 to 6 months}} & \multicolumn{2}{c}{\textbf{18 to 120 months}} \\ \hline
      geneset & E.S. & q value & E.S. & q value & E.S. & q value \\\hline
      Acetate syn. I & -0.122 & 0.555 & -0.194 & \color{blue}{\textbf{0.153}} & 0.127 & 0.602 \\
      Acetate syn. II & -0.22 & \color{blue}{\textbf{0.066}} & -0.342 & \color{blue}{\textbf{0.02}} & -0.095 & 0.92 \\
      Acetate syn. III & -0.21 & \color{blue}{\textbf{0.086}} & -0.31 & \color{blue}{\textbf{0.052}} & -0.086 & 0.92 \\
      Glutamate deg. I & -0.404 & \color{blue}{\textbf{0.18}} & -0.284 & 0.816 & -0.249 & 0.602 \\
      Glutamate syn. II & 0.095 & 0.602 & -0.134 & 0.98 & 0.242 & \color{red}{\textbf{0.047}} \\
      Menaquinone syn. I & -0.085 & 0.563 & -0.083 & 0.816 & -0.17 & \color{blue}{\textbf{0.182}} \\
      Propionate deg. I & -0.461 & \color{blue}{\textbf{0.13}} & -0.542 & \color{blue}{\textbf{0.02}} & -0.674 & \color{blue}{\textbf{0.041}} \\
      Propionate syn. I & -0.334 & 0.555 & -0.32 & 0.571 & -0.676 & \color{blue}{\textbf{0.023}} \\
      Tryptophan syn. & 0.057 & 0.974 & -0.072 & 0.816 & 0.119 & \color{red}{\textbf{0.041}} \\\hline\hline
    \end{tabular}
    \caption{\label{tab:fsea}Feature set enrichment analysis on neuroactive microbial gene sets.
    Red and blue text indicates significant (q value \textless 0.2) genesets with a positive or negative
    pseudomedian respectively.}
    \end{center}
\end{table}

@@

As with LMs, RF models for children under 6 months old
performed poorly (mean test-set correlation -0.12, mean RMSE 12.70),
but RFs were consistently able to learn the relationship between taxa
and cognitive function scores in children over 18 months of age (mean
test-set correlation 0.429, mean RMSE 17.29). For both age groups,
species that were important in RF overlapped with those that were
significant when testing the relationship with LMs (Figure
3A-B, Table S1A-B); except for \emph{Eubacterium ramulus}
all taxa that were significant using LMs belonged to a group
responsible for 80\% of the total importance in RF models
(Table 3, Figure 3A).
However, the rank-ordering of LM significance and RF importance
had substantive differences. For example, in children under 6 months,
only \emph{B. wexlerae} was significantly associated with cognitive function in LMs, but
was ranked 20th in importance for RF models.

>>>>

@@

% figures
\usepackage{wrapfig}
\usepackage{lscape}
\usepackage{rotating}
\usepackage{graphicx}
\graphicspath{assets}
\usepackage{float}
\floatstyle{plaintop}
\restylefloat{table}

@@

\begin{table}[!h]
    \begin{center}
    \begin{tabular}{|r|r|r|r|r|r|r|}
      \hline\hline
      \textbf{Age group} & \multicolumn{2}{c}{\textbf{0 to 120 months}} & \multicolumn{2}{c}{\textbf{0 to 6 months}} & \multicolumn{2}{c}{\textbf{18 to 120 months}} \\ \hline
      geneset & E.S. & q value & E.S. & q value & E.S. & q value \\\hline
      Acetate syn. I & -0.122 & 0.555 & -0.194 & \color{blue}{\textbf{0.153}} & 0.127 & 0.602 \\
      Acetate syn. II & -0.22 & \color{blue}{\textbf{0.066}} & -0.342 & \color{blue}{\textbf{0.02}} & -0.095 & 0.92 \\
      Acetate syn. III & -0.21 & \color{blue}{\textbf{0.086}} & -0.31 & \color{blue}{\textbf{0.052}} & -0.086 & 0.92 \\
      Glutamate deg. I & -0.404 & \color{blue}{\textbf{0.18}} & -0.284 & 0.816 & -0.249 & 0.602 \\
      Glutamate syn. II & 0.095 & 0.602 & -0.134 & 0.98 & 0.242 & \color{red}{\textbf{0.047}} \\
      Menaquinone syn. I & -0.085 & 0.563 & -0.083 & 0.816 & -0.17 & \color{blue}{\textbf{0.182}} \\
      Propionate deg. I & -0.461 & \color{blue}{\textbf{0.13}} & -0.542 & \color{blue}{\textbf{0.02}} & -0.674 & \color{blue}{\textbf{0.041}} \\
      Propionate syn. I & -0.334 & 0.555 & -0.32 & 0.571 & -0.676 & \color{blue}{\textbf{0.023}} \\
      Tryptophan syn. & 0.057 & 0.974 & -0.072 & 0.816 & 0.119 & \color{red}{\textbf{0.041}} \\\hline\hline
    \end{tabular}
    \caption{\label{tab:fsea}Feature set enrichment analysis on neuroactive microbial gene sets.
    Red and blue text indicates significant (q value \textless 0.2) genesets with a positive or negative
    pseudomedian, respectively.}
    \end{center}
\end{table}

@@

As with LMs, RF models for children under 6 months old
performed poorly (mean test-set correlation -0.12, mean RMSE 12.70),
but RFs were consistently able to learn the relationship between taxa
and cognitive function scores in children over 18 months of age (mean
test-set correlation 0.429, mean RMSE 17.29). For both age groups,
species that were important in RF overlapped with those that were
significant when testing the relationship with LMs (Figure
3A-B, Table S1A-B); except for \emph{Eubacterium ramulus},
all taxa that were significant using LMs belonged to a group
responsible for 80\% of the total importance in RF models
(Table 3, Figure 3A).
However, the rank-ordering of LM significance and RF importance
had substantive differences. For example, in children under 6 months,
only \emph{B. wexlerae} was significantly associated with cognitive function in LMs, but
was ranked 20th in importance for RF models.
